### 2016-10-10 lists_conversion_experience.rmd
got to infrastructure

# Introduction: the presentation of a presentation

@@

In this respect, the presentation was not really surprising or exceptional. Its ordinariness and prosaic character stands out as the most significant feature.  The presentation itself is a working example for an argument about our experience of predictive ordering.

@@

If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.

@@

# 1. The context of the presentation - Consumption and shopping
- presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities.  Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.

@@

- Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco. I found her presentation on shopping lists to be very practical since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways. 
Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers. [@Humby_1984]  
Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984]

@@

@@

# 2. The web and the shopping basket: how it was reconstituted in the 1990s
- online commerce -- transforming shopping; 
- personalization and recommendation as the killer apps for social media as a commercial process, but also in government etc.

@@

# 3. From the rules-based list to a 'relevancy model'

@@

# 4. Hack? Customers go straight to the checkout

@@

# 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?

@@

# 6. Gain: use good features to build models 
- what counts as a good feature? See Domingos on this
- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
- the model counts for surprisingly little -- at least, they all seem similar ... rather what counts is all that stuff on working with substitutions, similiarities and complementarities for each person.

@@

# 7. Sparsity: output 20 recommendations from 200,000 products

@@

# 8. Value: Balancing conversion vs spend
The value of an ordering

# 9. Metric: precision-at-K

@@

# 10. Experiment: A/B allocation

@@

Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommendation system. This is done in the name of statistical validation of the effects or the uplift of the model.

# 11. Infrastructure: realtime engine serves up top 5
Only in certain settings are things realtime; or only in limited ways;
Realtime has its own economies and cultures

# 12. Similarity, substitution and complementarity
- interesting contrasts here between the relations of things in a list
- whereas the `arule` approach compares lists looking for similiarities and ways of completing lists, the more recent approaches look for micrological propensities in the list that relate to a particular time and place, amidst a continuously differentiable pathways between things ...

# 13. Complementarity drives conversion 
- the problem of action

# Conclusion
- lists have complex temporal structure
- listing practices very closely entangled with problems of social order
- lists undergoing a 'conversion' experience as they become more probabilistic

>>>>

## Introduction: the presentation of a presentation

@@

In this respect, the presentation was not really surprising or exceptional. Its ordinariness and prosaic character stands out as the most significant feature.  The presentation itself is a working example for an argument about our experience of predictive ordering. In this paper, I re-present the presentation using a mixture of theoretical and practical resources. The theory is largely recent media theory, and the practical resources are that of coding and predictive models.

@@

If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or in terms of the predictive conditions of contemporary experience [@Hansen_2015a] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.

@@

## 1. The context of the presentation - Consumption and shopping

I heard this presentation at an academic event with speakers from industry and government on statistics and operations research in industry. Speakers came from Shell, Disney, GCHQ, and various universities.  Because I've done a lot of work on machine learning and data science more generally, I found the sessions both familiar and occasionally exciting.

@@

Shreena Patel works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/)  focusing on online grocery shopping at Tesco. I found her presentation on shopping lists to be  practical and evocative since it was focused quite a lot on implementation of recommendations, and symptomatic in various ways.

Tesco is the largest supermarket chain in UK, famous for Clubcard starting in 1991; often cited as a leading early example of Big Data Analytics. 
DunnHumby -- Edwina Dunn and Clive Humby -- are famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers. [@Humby_1984]  
Clive Humby's academic work is hard to track down, but an early paper of his given at the Conference of Your Operational Researchers in Nottingham in 1984 suggests the direction that he and later Tesco would take in constructing lists [@Humby_1984]

@@

While Tesco has been very successful in data-mining its customers, the online shopping situation is a bit different. It is no longer organised around campaigns involving special offers or redemption of points, but on continuous up-selling and cross-selling.

@@

## 2. The web and the shopping basket: how it was reconstituted in the 1990s

![People add to list](figure/online_lists_in_practice.png)

- online commerce -- transforming shopping; 
- personalization and recommendation as the killer apps for social media as a commercial process, but also in government etc.

@@

## 3. From the rules-based list to a 'relevancy model'

@@

## 4. Hack? Customers go straight to the checkout

@@

## 5. Data: 52 weeks of data for each customer: how is order constructed -- from history?

@@

## 6. Gain: use good features to build models 
- what counts as a good feature? See Domingos on this
- move from Dunnhumby's ClubCard focus to much more general world of predictive models (e.g. Facebook use a million models, etc.)
- the model counts for surprisingly little -- at least, they all seem similar ... rather what counts is all that stuff on working with substitutions, similiarities and complementarities for each person.

@@

## 7. Sparsity: output 20 recommendations from 200,000 products

@@

## 8. Value: Balancing conversion vs spend
The value of an ordering

## 9. Metric: precision-at-K

@@

## 10. Experimental settings: A/B allocation and uplift

@@

Random allocation of customers to the four categories (Test A, Test B, Control A, Control B) adds another layer of probability to the recommendation system. This is done in the name of statistical validation of the effects or the uplift of the model. It also suggests that the effects of the model cannot be known in advance -- they can only be observed experimentally. Does this experimentality of the list point more to the propensities that Hansen says saturate experience?

It is perhaps significant that this random allocation of customers to control and test groups occurs without any relation to the particular profile or propensities of the customer. It seeks to statistically validate the effects -- the *uplift* -- of the model on  conversion events by allowing the effects of the model on what people do to be measured. The lift or uplift refers to a change in events associated with the same groups of people. This effectively is an experiment in inhabiting two different worlds.  This sets up a feedback look between the predictive model, the logistic regression, and the operation of the recommendation system.[^12] Once the experimental uplift methodology is in place, the recommendation system can constantly modify the recommendation list in light of subsequent data on the uplift. Predictions, then, are not bound by fixed parameters of the model.

[^12]: See [@Lo_2002] for an introduction to the construction of experimental set ups for uplift models. Market researchers and data miners seem to have started adopting a statistically grounded experimental predictive practice sometimes around 2000. See for instance [@Almquist_2001]  Response modelling, propensity modelling and uplift modelling all seek to identify associations between 'treatments' or interventions and the 'Responders', the people  affected by the treatment.  Interestingly, these

## 11. Infrastructure: realtime engine serves up top 5

![Realtime infrastructure and its limits](~/lists/figure/hadoop.png)

While much discussion of recent media has highlighted its realtime operation, Shreena's description of the infrastructure that implements the new model suggested a different temporal patterning. While raw data from Tesco Online feeds into the model every hour, it seems that recommendation list for each customer is only generated once a week. This makes sense. Customers only shop online every few days at the most, and in some cases, only every few weeks. To continually update the top 200 recommendations for every customer would be computationally expensive.

Shreena did not say much else about the infrastructure but the 200 recommendations generated each week are certainly never seen by the customers. The realtime engine that actually generates results shown on the Tesco Online website only shows 5 after 'applying business rules.' Again, this might be seen as corroborating Hansen and Badiou's insistence on the necessarily limited grasp of any prediction system on the situation it encounters. Even the new model, situated amidst its experimental setup cannot fully encompass the surplus of sensibility it encounters.

The brief mention of specific infrastructural elements such as `hadoop` is, however, significant. It points to the temporality of infrastructures themselves as a component of list practices. The possibility of adjusting the recommendations for every customer weekly derives from an infrastructure capable of collecting data, assimilating that data to predictive model able to generate recommendations that are relevant to individuals amongst 200,000 products.

## 12. Similarity, substitution and complementarity

- interesting contrasts here between the relations of things in a list
- whereas the `arule` approach compares lists looking for similiarities and ways of completing lists, the more recent approaches look for micrological propensities in the list that relate to a particular time and place, amidst a continuously differentiable pathways between things ... 
- how do things associate with each other in complementarities, substitutions, etc.

## 13. Complementarity drives conversion 
- the problem of action

## Conclusion
- lists have complex temporal-material structure, since they mix prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the surplus of sensibility associated with commodity production. 
- listing practices very closely entangled with problems of social order under predictive conditions or amidst predictive operations. 
- lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset. Amidst prediction they become constitutively incomplete.

## References
